Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-05-17 SIR_manuscript.tex
minor changes in the abstract

The effect of testing processes, including testing and test reporting, on an epidemic dynamics, involving infection and recovery, can be studied at the individual level or the community level (e.g., nursing homes, long-term-care facilities, etc.). 
Gaining insights to determine the sensitivity of the epidemic dynamics with respect to the testing processes will depend on underlying factors including the level of focus (individual or community), assumptions (model), and the interplay between these factors. 
In particular, the fast testing and test reporting may be beneficial at the community-level, supported by many studies, as it gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing. However, the potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified. At the individual level, this advantage can manifest in the following sense: individuals awaiting their test results or who have tested positive may partially or fully self-isolate, thus reducing or eliminating their potential in the transmission process.
In this paper, we investigated this individual-level effect of testing processes on the epidemic dynamics by developing a SIR-type model.
Although the model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures. The realistic components of the model include \percap testing intensity, test sensitivity and specificity, rate of test return, and isolation efficacy in reduction of the probability of transmission. The novel component is the compartment-specific relative testing weights, which reflect the testing strategies-- surveillance, diagnosis, or control.
Here, we compare two testing strategies, random vs. targeted, and concluded that the targeted testing strategy is more effective in the sense that achieving reduction of $\Rnum$ can occure in a lower range of testing intensity and longer test return time relative to the random testing. Furthermore, we show that increasing \percap testing intensity and reducing the test return time would be beneficial on the dynamics of an epidemic in general but there are exception cases. In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the \percap testing intensity and the rate of test return.

>>>>

The effect of testing processes, including testing and test reporting, on an epidemic dynamics, involving infection and recovery, can be studied at the individual level or the community level (e.g., nursing homes, long-term-care facilities, etc.). 
Gaining insights to determine the sensitivity of the epidemic dynamics with respect to the testing processes will depend on underlying factors including the level of focus (individual or community), assumptions (model), and the interplay between these factors. 
In particular, the fast testing and test reporting may be beneficial at the community-level, supported by many studies, as it gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing. However, the potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified. At the individual level, this advantage can manifest in the following sense: individuals awaiting their test results or who have tested positive may partially or fully self-isolate, thus reducing or eliminating their potential in the transmission process.
In this paper, we investigated this individual-level effect of testing processes on the epidemic dynamics by developing a SIR-type model.
Although the model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures. The realistic components of the model include \percap testing intensity, test sensitivity and specificity, rate of test return, and isolation efficacy in reduction of the probability of transmission. The novel component is the compartment-specific relative testing weights, which reflect the testing strategies-- surveillance, diagnosis, or control. Here, we compare two testing strategies, random vs. targeted, and concluded that increasing testing “focus” from random to targeted always decreases \Rnum. Further, the following processes always decreases $\Rnum$; increasing the isolation efficacy parameters for tested and confirmed individuals, higher testing intensity if testing is random or testing intensity is small, and a higher rate of test return if the isolation efficacy for tested-awaiting individuals is low.
